---
title: "vacc_tests"
format: html
---

## Import data

```{r}
#| echo: false

# Import data 
library(readr)
cleanvac <- read_csv("project841_7.csv")

# variable names 
library(tidyverse)
tibble(cleanvac)


```

```{r}

# stats of aggregate beliefs and exp scores
summary(cleanvac$beliefs1)
summary(cleanvac$exp1)


# check normality assumption
# both scores fail to meet normality assumption
shapiro.test(cleanvac$beliefs1)
shapiro.test(cleanvac$exp1)

# transform scores 
# range = [0,1]
# interpret as proportion of questions answered favorably by diving out of total questions 


# belief score, max questions = 10
cleanvac$t_beliefs <- (cleanvac$beliefs1)/10

# experience score, max questions = 14
cleanvac$t_exp   <- (cleanvac$exp1)/14

```

```{r}
# recheck stats and normality assumption
# still fails to meet normality assumption
summary(cleanvac$t_beliefs)
summary(cleanvac$t_exp)

shapiro.test(cleanvac$t_beliefs)
shapiro.test(cleanvac$t_exp)

hist(cleanvac$beliefs1)
hist(cleanvac$t_beliefs)

hist(cleanvac$exp1)
hist(cleanvac$t_exp)

```



### Use non-parametric tests: 
# 1. Wilcoxon rank sum test (vs. 2 groups, binary)


```{r}
table(cleanvac$caregiver_sex)
class(cleanvac$caregiver_sex)

library(coin)
library(rstatix)
# Wilcoxon Rank Sum test, binary 
wilcox.test(t_beliefs ~ caregiver_sex, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ caregiver_sex,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)

wilcox.test(t_beliefs ~ bednet, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ bednet,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_beliefs ~ inpatient, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ caregiver_sex,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_beliefs ~ child_sex, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ child_sex,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_beliefs ~ hbv_test, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ hbv_test,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_beliefs ~ hbv_res, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ hbv_res,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_beliefs ~ dbs, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ dbs,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)

wilcox.test(t_exp ~ caregiver_sex, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ caregiver_sex,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ bednet, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ bednet,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ inpatient, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ inpatient,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ child_sex, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ child_sex,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ hbv_test, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ hbv_test,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ hbv_res, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ hbv_res,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ dbs, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ dbs,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)

wilcox.test(t_beliefs ~ card, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_beliefs ~ card,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)
wilcox.test(t_exp ~ card, data=cleanvac)
wilcox_effsize(
  cleanvac,
  t_exp ~ card,
  ref.group = "0",
  paired = FALSE,
  alternative = "two.sided"
)




```


### Wilcoxon boxplots, sig. associations
```{r}
#| echo: false
library(ggstatsplot)


# belief plots
ggbetweenstats(
  data = cleanvac,
  x = bednet,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)


ggbetweenstats(
  data = cleanvac,
  x = hbv_test,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)



ggbetweenstats(
  data = cleanvac,
  x = inpatient,
  y = t_exp,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)


ggbetweenstats(
  data = cleanvac,
  x = card,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = card,
  y = t_exp,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)




```



# 2. Kruskal-Wallis test (vs. 3 or more groups, nominal)

```{r}
#| echo: false
# Kruskal-Wallis, 3 or more groups, nominal

kruskal.test(t_beliefs ~ relationship, data=cleanvac)
kruskal.test(t_beliefs ~ marital, data=cleanvac)
kruskal.test(t_beliefs ~ vacc_where, data=cleanvac)
kruskal.test(t_beliefs ~ birthplace2, data=cleanvac)
kruskal.test(t_beliefs ~ birth_status, data=cleanvac)
kruskal.test(t_beliefs ~ muac_color, data=cleanvac)
kruskal.test(t_beliefs ~ education, data=cleanvac)

kruskal.test(t_exp ~ relationship, data=cleanvac)
kruskal.test(t_exp ~ marital, data=cleanvac)
kruskal.test(t_exp ~ vacc_where, data=cleanvac)
kruskal.test(t_exp ~ birthplace2, data=cleanvac)
kruskal.test(t_exp ~ birth_status, data=cleanvac)
kruskal.test(t_exp ~ muac_color, data=cleanvac)
kruskal.test(t_exp ~ education, data=cleanvac)

```

## Dunn test

```{r}
# Multiple comparison tests after sig. association in kruskal-wallis

options(tibble.print_max = Inf) 

library(rstatix)
dunn_test(data=cleanvac, t_beliefs ~ vacc_where, p.adjust.method = "holm", detailed = TRUE)
dunn_test(data=cleanvac, t_beliefs ~ birthplace2, p.adjust.method = "holm", detailed = TRUE)
dunn_test(data=cleanvac, t_beliefs ~ muac_color, p.adjust.method = "holm", detailed = TRUE)
dunn_test(data=cleanvac, t_beliefs ~ education, p.adjust.method = "holm", detailed = TRUE)


dunn_test(data=cleanvac, t_exp ~ vacc_where, p.adjust.method = "holm", detailed = TRUE)
dunn_test(data=cleanvac, t_exp ~ marital, p.adjust.method = "holm", detailed = TRUE)
dunn_test(data=cleanvac, t_exp ~ education, p.adjust.method = "holm", detailed = TRUE)
```

## K-W plots

```{r}
#| echo: false
# belief plots
library(ggstatsplot)
ggbetweenstats(
  data = cleanvac,
  x = vacc_where,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = birthplace2,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = muac_color,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = education,
  y = t_beliefs,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

# exp plots
ggbetweenstats(
  data = cleanvac,
  x = vacc_where,
  y = t_exp,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = marital,
  y = t_exp,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = education,
  y = t_exp,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)


```

### Untransformed scores
```{r}
#| echo: false
# untransformed scores
# belief plots
library(ggstatsplot)
ggbetweenstats(
  data = cleanvac,
  x = vacc_where,
  y = beliefs1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = birthplace2,
  y = beliefs1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = muac_color,
  y = beliefs1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = education,
  y = beliefs1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

# exp plots
ggbetweenstats(
  data = cleanvac,
  x = vacc_where,
  y = exp1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = marital,
  y = exp1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)

ggbetweenstats(
  data = cleanvac,
  x = education,
  y = exp1,
  type = "nonparametric", 
  plot.type = "box",
  pairwise.comparisons = TRUE,
  pairwise.display = "significant",
  centrality.plotting = TRUE,
  bf.message = FALSE
)
```


# 3. Kendall's tau (vs. continuous)


```{r}

# Kendall's tau corr, continuous

cor.test(cleanvac$t_beliefs, cleanvac$t_exp, method = "kendall")
cor.test(cleanvac$t_beliefs, cleanvac$age_months, method = "kendall")
cor.test(cleanvac$t_beliefs, cleanvac$oldest_child, method = "kendall")
cor.test(cleanvac$t_beliefs, cleanvac$muac, method = "kendall")
cor.test(cleanvac$t_beliefs, cleanvac$children, method = "kendall")
cor.test(cleanvac$t_beliefs, cleanvac$birthplace, method = "kendall")

cor.test(cleanvac$t_exp, cleanvac$age_months, method = "kendall")
cor.test(cleanvac$t_exp, cleanvac$oldest_child, method = "kendall")
cor.test(cleanvac$t_exp, cleanvac$muac, method = "kendall")
cor.test(cleanvac$t_exp, cleanvac$children, method = "kendall")
cor.test(cleanvac$t_exp, cleanvac$birthplace, method = "kendall")

```

### Corr plots, sig. associations

```{r}
#| echo: false

library(ggstatsplot)
ggscatterstats(data=cleanvac, t_beliefs, muac, type="nonparametric", conf.level=0.95)

ggscatterstats(data=cleanvac, t_exp, muac, type="nonparametric", conf.level=0.95)


ggscatterstats(data=cleanvac, t_exp, oldest_child, type="nonparametric", conf.level=0.95)


ggscatterstats(data=cleanvac, t_exp, children, type="nonparametric", conf.level=0.95)



 
```

# Brief Results

Caregivers who brought their child to government health facilities 2, 3, or Immunization Program Visits to receive vaccinations over private healthcare were associated with a higher aggregate belief score (Table S1). Caregivers who did not complete school compared to those who completed primary or secondary school, and children born in private healthcare facilities over government health facilities or at home were associated with a lower aggregate belief score (Table S1). Children who slept under a bed net in the previous night were associated with a higher aggregate belief score, although with minimal effect (Cohen’s d = 0.089, p-value = <0.001, Table 3).

Caregivers taking care of more children were associated with a marginally higher aggregate experience score (Kendall’s tau = 0.070, p-value = < 0.001, Table 3). Caregivers who brought their child to government health facility 2 over facility 3 to receive vaccinations were associated with a higher aggregate experience score (Table S1). In contrast to belief scores, caregivers who did not complete school compared to those who completed primary and secondary school, and married caregivers over those unmarried were associated with a higher aggregate experience score (Table S1). Children who stayed overnight in a hospital or clinic at least once in their lifetime were associated with a higher aggregate experience score, although with small effect (Cohen’s d = 0.131, p-value = <0.001, Table 3).


